### 2014-02-25 bullet_groups.tex
Article

\documentclass[12pt,preprint]{aastex}
% To highlight comments 
\usepackage{color}
\definecolor{red}{rgb}{1,0.0,0.0}
\newcommand{\red}{\color{red}}

\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{graphics}
\usepackage{epsfig}  
\usepackage{float}
%\bibliographystyle{apj}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}

\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\bullb}{MACS J0025.4-1222}
\newcommand{\bulla}{1E0657---56} 
\shorttitle{Bullet Groups}
\shortauthors{Fern\'andez-Trincado et al.}

\begin{document}

\title{Bullet Groups in the Bolshoi Cosmological Simulation}
\author{J. G. Fern\'andez-Trincado$^{1,2,3}$, J. E. Forero-Romero$^1$ and T. Vergudo$^3$}
\affil{$^1$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\
       $^2$ Institute Utinam, CNRS UMR6213, Universit\'e de Franche-Comt\'e, OSU THETA de Franche-Comt\'e-Bourgogne, Besan\c{c}on, France\\
       $^3$ Centro de Investigaciones de Astronom\'ia, AP 264, M\'erida 5101-A, Venezuela}

\begin{abstract}

We estimate the expected distribution of displacements between the host halo of dark matter and the 
substructure of dark matter most massive of the host, using a halo catalog from the Bolshoi simulation detected 
with the Bound Density Maximun (BDM) algorithm. We found that the probability of finding a system similar to the Bullet
Group with displacements of $\sim$400 kpch$^{-1}$ is between 40\% for z$=0$ and 60\% for z$=1$ and $\Lambda$CDM standard 
model predicts this scenario and is consistent with recently observations of the object SL2S J08544-0121 that is 
a gravitational lens found in the SL2S and located at z$=0.35$.

\end{abstract}

\begin{keywords}
{cosmology: theory -- dark matter} 
\end{keywords}

\section{Introduction}

This paper is organized as the follows: In Section
\ref{simulation} we present the simulations.

\section{Simulation}
\label{simulation}

The analysis presented in this paper uses mainly the Bolshoi (250 Mpch$^{-1}$ simulation box, 1 kpch$^{-1}$ resolution) Database simulation described in
\citet{2011ApJ...740..102K}. The simulation follows the evolution of 8.6 billion particle cosmological N-body simulation
from z=80 to z=0 in a comoving cube of $250$ Mpch$^{-1}$
on a side. The cosmology used  corresponds to  the spatially flat
concordance model with the
following parameters:  the density parameter for matter (dark matter$+$baryons) $\Omega_m=0.27$, the density parameter for baryonic matter
 $\Omega_b=0.0469$, the density parameter for dark energy $\Omega_{\Lambda}=0.73$, the Hubble
parameter $h=0.7$, the normalization of the Power spectrum $n=0.95$ and
the amplitude of mass density fluctuation (at redshift z$=$0) $\sigma_8=0.82$.  The number of particles used for each of
the DM component was $2048^3$, resulting in a mass resolution of $1.35 \times
10^8$ M$_{\odot}$h$^{-1}$.

\section{Cumulative Probability Distributions}
\label{cumulative distribution}

We selected four snapshot of the simulation for four different redshifts (z$=0$, z$=0.25$, z$=0.5$ and z$=1$) based in the 
distribution of redshift in the Figure 1 by \citet{verdugo}, and 
for each sample in redshift we selected the host halo with circular velocities greater than 300 kms$^{-1}$ and was split in two 
principal groups: The first group correspond to host halo with circular velocities between 300 kms$^{-1}$ to 700 kms${^{-1}}$
with mass in the range of $10^{12}$ M$_\odot{}$ to $10^{14}$ M$_\odot{}$ in the range of mass of the
Bullet Groups, and second group correspond to host halo with circular velocities greater than 
700 kms$^{-1}$ with mass $\geq10^{14}$ M$_\odot{}$, in the range of mass of the Bullet Clusters, in 
the Table \ref{table1} is shown the two groups selected in this work. For each group, we 
classified the corresponding substructures most massive and associated with the corresponding 
host halo. The configuration of this system is shown in 
Figure \ref{configuration}, where you can distinguish the host halo and substructure for a particular configuration in z$=0$. In 
this work, we estimate the expected distribution of displacements ($d_{real,(X,Y)}$) in the projection 2-D that can estimate by 
observations, these displacements correspond to the separation between the minimal potential of the host halo and the 
minimal potential of the substructure, both are dark matter distributions.\\

\begin{table*}
\begin{center}
\begin{tabular}{ccccccc}\hline\hline
Sample       & Minimum Mass           & Maximum Mass          & \# Halos  & \# Halos  & \# Halos & \# Halos \\
             & M$_{\odot}$h$^{-1}$    & M$_{\odot}$h$^{-1}$   &   $z=0$   & $z=0.25$  & $z=0.5$  & $z=1$  \\\hline
Host Halo    & 0.35$\times{}10^{14}$  & 1.09$\times{}10^{14}$ &    400    &   362     &   310    &  192   \\
Substructure & 0.38$\times{}10^{11}$  & 0.30$\times{}10^{14}$ &    400    &   362     &   310    &  192   \\\hline
\end{tabular} 
\caption{Mass ranges for the two groups selected in this work in different redshifts.}
\label{table1}
\end{center}
\end{table*}

In order to explore the distribution of displacements expected in the observational, we define a new parameter given by:\\

\begin{equation}
 \frac{\nu_{circ,sub}}{\nu_{circ,halo}}=0.5
\end{equation}

Figure \ref{newparameter} shows the scatter plot of the parameter $\left(\frac{\nu_{circ,sub}}{\nu_{circ,halo}}\right)$ vs the 
displacement ($d_{2d,(X,Y)}$) between the host halo and the substructure and for different redshifts. This parameter is 
consistent with the observations. The red star simbol in the Figure \ref{newparameter} is equal to 0.54 corresponding to 
the fraction in velocity dispersion in the line-of-sigth of the group SL2S SJ08544-0121 ($\sigma_{host,halo}=341^{+43}_{-109}$ kms$^{-1}$ and $\sigma_{substructure}=185^{+30}_{-62}$), reported by \citet{2013A&A...552A..80M}.
Based in the parameter $\left(\frac{\nu_{circ,sub}}{\nu_{circ,halo}}\right)>0.5$, we estimate the expected distribution of
displacements and this is shown in the Figure \ref{displacements} for the two groups classified in this work.

\begin{figure*}
\begin{center}
\includegraphics[width=0.90\textwidth]{Bullet_Group.eps}
\end{center}
\caption{Configuration for a host halo (dashed circle) and sub-structure (black circle) in the sample. } 
\label{configuration}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{figure_7_2.eps}
\includegraphics[width=1.0\textwidth]{figure_7_3.eps}
\end{center}
\caption{Scatter $\left(\frac{\nu_{circ,sub}}{\nu_{circ,halo}}\right)$ vs $d_{2d,(X,Y)}$ for different redshifts. The 
horizontal dashed line correpond to $\left(\frac{\nu_{circ,sub}}{\nu_{circ,halo}}\right)=0.5$ and the red star simbol correspond
to $\left(\frac{\nu_{circ,sub}}{\nu_{circ,halo}}\right)=0.54$ for SL2S SJ08544-0121 in \citet{2013A&A...552A..80M}. The 
four top panels are the sample with circular velocities $>700$ kms$^{-1}$ and the four down panels are the sample with 
circular velocities between 300 km s$^{-1}$ to 700 kms$^{-1}$.}
\label{newparameter}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.1\textwidth]{figure_7_1.eps}
\end{center}
\caption{Cumulative distribution (P$>d_{2d,(X,Y)}$) of displacements for the projection (X,Y). {\bf Left panel:} Sample with $\nu_{max}>700$ kms$^{-1}$ 
for different redshifts. The vertical dashed line, correspond to the separation between dark matter to dark matter estimate
in this work as the double of separation between the collisional gas and dark matter of 124$\pm$20 kpc reported
by \citet{gastaldello} for the group SL2S J08544-0121. 
{\bf Right panel:} Sample with $300 $ kms$^{-1} <\nu_{max}<700 $ kms$^{-1}$ for different redshifts.} 
\label{displacements}
\end{figure*}

\subsection{Number expected of Bullet Groups in the sample}

In order of estimate the number of Bullet Groups expected in the Bolshoi Cosmological Simulation, we 
define the configuration of this system as one where the substructure is coming out of the host halo ($cos(\theta)>0.5$). For this 
we make the scalar product between the velocity vector and position vector that defines the separation between 
 the host halo and the substructure, as shown below: \\

 \begin{equation}
  cos(\theta)=\frac{\vec{\nu{}}\cdotp{}\vec{r}}{\left\|\nu{}\right\| \left\|r\right\|}
 \end{equation}

where, $\vec{\nu{}}=\vec{\nu}_{sub}-\vec{\nu}_{halo}$ and $\vec{r}=\vec{r}_{sub}-\vec{r}_{halo}$, in the Figure \ref{cos_theta}
is shown the scatter plot of $cos(\theta{})$ vs $X_{off,new}=\frac{d_{2d,(X,Y)}}{R_{virial,halo}}=\frac{d_{real,(X,Y)}}{R_{virial,halo}}$.

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{figure_8_1_nu=0.5_300kms_700kms.eps}
\end{center}
\caption{Scatter of $cos(\theta)$ vs $X_{off,new}$. The red dashed line correpond to the limit for $cos(\theta{})>0.5$,  
where the substructure is emerging from the host halo and  $\left(\frac{\nu_{circ,sub}}{\nu_{circ,halo}}\right)\geq0.5$, 
cicular velocities $<700$ kms$^{-1}$ and different redshifts.} 
\label{cos_theta}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{figure_8_3_nu=0.5_300kms_700kms.eps}
\end{center}
\caption{Cumulative distribution of $d_{real, (X,Y)}$ for $cos(\theta)>0.5$, $\left(\frac{\nu_{circ,sub}}{\nu_{circ,halo}}\right)\geq0.5$, 
cicular velocities $<700$ kms$^{-1}$ and different redshifts.} 
\label{cumulative_cos}
\end{figure*}

\section{Summary}
\label{summary}

\section{Conclusions}
\label{conclusions}

\section*{Acknowledgements}

The CosmoSim database used in this paper is a service by the Leibniz-Institute for Astrophysics Potsdam (AIP). The 
BolshoiP simulation was performed within the Bolshoi project of the University of California High-Performance 
AstroComputing Center (UC-HIPACC) and was run at the NASA Ames Research Center.

\begin{thebibliography}{29}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{{Klypin} {et~al.}(2011)}]{2011ApJ...740..102K}
{Klypin}, A.~A. and {Trujillo-Gomez}, S. and {Primack}, J.} 2011, \apj, 102

\bibitem[{{Verdugo} {et~al.}(2014)}]{verdugo}
{{Vedugo}, T., {Motta}, V., {Foex}, G., {Forero-Romero}, J. E., {Mu\~noz}, R. P., {Pello}, R., 
{Limousin}, M., {Gavazzi}, R., {More}, A., {Cabanac}, R., {Soucail}, G., {Blakeslee}, J. P., 
{Mej\'ia}, A., {Magris}, G. and {Fern\'andez-Trincado}, J. G.} 2014, \apj, in preparation

\bibitem[{{Gastaldello} {et~al.}(2014)}]{gastaldello}
{{Gastaldello}}, F., et al. 2014, \mnras, in preparation

\bibitem[{{Mu\~noz} {et~al.}(2013)}]{2013A&A...552A..80M}
{{Mu{\~n}oz}, R.~P. and {Motta}, V. and {Verdugo}, T. and {Garrido}, F. and 
{Limousin}, M. and {Padilla}, N. and {Fo{\"e}x}, G. and {Cabanac}, R. and 
{Gavazzi}, R. and {Barrientos}, L.~F. and {Richard}, J.} 2013, \aap, A80

%%%%%%%%%%%%%%%%%%%%%%

%\bibitem[{{Tormen}(1997)}]{1997MNRAS.290..411T}
%{Tormen}, G. 1997, \mnras, 290, 411

%\bibitem[{{Tormen} {et~al.}(2004){Tormen}, {Moscardini}, \&
%  {Yoshida}}]{2004MNRAS.350.1397T}
%{Tormen}, G., {Moscardini}, L., \& {Yoshida}, N. 2004, \mnras, 350, 1397

%\bibitem[{{Vitvitska} {et~al.}(2002){Vitvitska}, {Klypin}, {Kravtsov},
%  {Wechsler}, {Primack}, \& {Bullock}}]{2002ApJ...581..799V}
%{Vitvitska}, M., {Klypin}, A.~A., {Kravtsov}, A.~V., {Wechsler}, R.~H.,
%  {Primack}, J.~R., \& {Bullock}, J.~S. 2002, \apj, 581, 799

%\bibitem[{{Yepes} {et~al.}(2007){Yepes}, {Sevilla}, {Gottl{\"o}ber}, \&
%  {Silk}}]{2007ApJ...666L..61Y}
%{Yepes}, G., {Sevilla}, R., {Gottl{\"o}ber}, S., \& {Silk}, J. 2007, \apjl,
%  666, L61

\end{thebibliography}

\end{document}