### 2014-03-20 paper/ms.tex
new version

\documentclass{emulateapj}
%\documentclass{aastex}
\submitted{{\it Submitted for publication in ApJL}}
\usepackage{multirow,color,wrapfig,ulem}
\usepackage {graphicx}

\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{graphics}
\usepackage{epsfig}  
\usepackage{float}
\bibliographystyle{apj}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}

\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}
\newcommand{\bullb}{MACS J0025.4-1222}
\newcommand{\bulla}{1E0657---56} 
\newcommand{\bullg}{SL2S J08544-0121}
\shorttitle{Bullet Groups}
\shortauthors{Fern\'andez-Trincado et al.}

\begin{document}

\title{The abundance of Bullet-Groups in $\Lambda$CDM}
\author{J. G. Fern\'andez-Trincado$^{1,2,3}$, J. E. Forero-Romero$^1$,
  T. Verdugo$^3$, G. Foex$^4$ and V. Motta$^4$} 
\affil{$^1$ Departamento de F\'{i}sica, Universidad de los Andes,
  Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\ 
       $^2$ Institute Utinam, CNRS UMR6213, Universit\'e de
  Franche-Comt\'e, OSU THETA de Franche-Comt\'e-Bourgogne,
  Besan\c{c}on, France\\ 
       $^3$ Centro de Investigaciones de Astronom\'ia, AP 264,
  M\'erida 5101-A, Venezuela\\
  $^4$ Instituto de F\'isica y Astronom\'ia, Universidad de
  Valpara\'iso, Avda. Gran Breta\~na 1111, Playa Ancha, Valpara\'iso
  2360102, Chile  
}
\email{jfernandezt87@gmail.com}
\email{je.forero@uniandes.edu.co}
\begin{abstract}

We estimate the expected distribution of displacements between the two
dominant dark matter density peaks and between baryons and dark matter
in halos simulated within a full cosmological context in $\Lambda$CDM.
We use as a benchmark the observation of \bullg, which is the lowest mass system observed so far
featuring a bi-modal dark matter distribution with a dislocated
baryonic component. We extend to lower masses previous computational results by
studying two halo samples: groups and clusters. We find that $(50 \pm
10)\%$ of the dark matter halos with circular velocities in the range
$300\kms$ to $700\kms$ (groups) show multimodal morphologies with
displacements between the dark matter clumps equal or larger than
$186 \pm 30$\hkpc as observed in \bullg. For dark matter halos with
circular velocities larger than $700\kms$ (clusters) this fraction
rises to $(70 \pm 10)\%$. Using the same simulation we estimate the
dark matter-to-baryon spatial separation and find that $0.1$ to
$1.0\%$ of the groups should present separations equal or
larger than $87\pm 14$\hkpc corresponding to our observational
benchmark; for clusters this fraction rises to $(7\pm 3)\%$, consistent
with previous studies of dark matter to baryon separations. In terms
of absolute number densities we find that the expected value for
groups similar to \bullg\ is three times larger than the expected value
for clusters. This extension of the theoretical predictions and
observational results towards low mass bi-modal dark matter
configurations opens up the possibility for a new statistical test of
$\Lambda$CDM.   
\end{abstract}

\keywords{dark matter, galaxies: clusters: individual (SL2S
  J08544-0121), galaxies: interactions, methods: numerical}

\section{Introduction}

The Bullet Cluster (\bulla) provided a new kind of observational
evidence for the existence of dark matter
\citep{Markevitch2004,Clowe2006}. Since then it has been used to test
the Cold Dark Matter (CDM) paradigm itself by quantifying different
aspects such as the expected displacement between the dominant dark
matter and baryonic component in a $\Lambda$CDM Universe
\citep{ForeroRomero2010} or the substructure velocity required to
produce such displacement \citep{Hayashi2006, Springel2007,
  Thompson2012}. It has also been used to constrain the dark matter
particle self-interaction cross section and to explore possible
extensions to the concordance cosmological model
\citep{Farrar2007,Lee2010,Lee2012}.

Since then, other examples of Bullet-like systems have been found; 
MACS J0025.4-1222 \citep{Bradac2008}, Abell 2744 \citep{Merten2011},
DLSCL J0916.2+2951 \citep{Dawson2012}, ZwCl 1234.0+02916
\citep{Dahle2013}. Recently \citep{Gastaldello} observed a DM-baryon
displacement of $87\pm 14$\hkpc and a DM-DM separation of $186\pm
30$\hkpc in \bullg, a group-like system with a total
mass $2.4\pm 0.6 \times 10^{14}$\Msun found in the the Strong Lensing
Legacy Survey (SL2S) sample \citep{Cabanac2007,More2012}.

Groups around this mass are $\sim10$ times more abundant than clusters
in the mass range of the Bullet Cluster $>10^{15}$\hMsun. This 
should open up the possibility of finding Bullet Groups in large
numbers to test $\Lambda$CDM. However, a larger abundance of small
mass systems has to be weighted by the probability of being in a
merger and  presenting a large displacement between the DM and
baryonic components.  These two conditions (merger rates, maximum
possible displacement) are a funcion of DM halo mass in $\Lambda$CDM
cosmologies.  A detailed statistical study to estimate the abundance
of Bullets has been performed for clusters but not for lower mass
systems \citep{ForeroRomero2010}.

In this Letter we extend such study for systems in the group mass
scale. We measure the abundance of DM systems with with a multi-modal
morphology (large DM-DM displacements) and estimate the amount of
systems with a Bullet configuration  (large DM-baryon
displacements). To this end we use a high resolution N-body
cosmological simulation (Bolshoi) that allows us to find  multi-modal
dark matter distributions in hosts with circular velocities larger
than $300$\kms ($\sim 1.0\times 10^{13}\hMsun$).

This Letter is organized as follows. In Section
\ref{sec:simulation} we present the simulation and the halo catalogs
used in this work. We continue in Section \ref{sec:setup} with the
geometry of the problem at hand and the measurement setup. Next in
Section \ref{sec:results} we present our results and observational
perspectives to finally conclude in Section \ref{sec:conclusions}.

\section{Simulation, halo catalogs and pairs}
\label{sec:simulation}

We use the Bolshoi Run, a cosmological DM only simulation over a cubic
volume of 250\hMpc comoving on a side. The simulation uses the ART
code \citep{Kravtsov1997} to follow the evolution of a dark matter
density field from $z=80$ to $z=0$ sampled with $2048^3$  particles. The
cosmology used  corresponds to  the spatially flat concordance model
with the following parameters:  the density parameter for matter (dark
matter and baryons) $\Omega_m=0.27$, the density parameter for baryonic matter
$\Omega_b=0.0469$, the density parameter for dark energy
$\Omega_{\Lambda}=0.73$, the Hubble parameter $h=0.7$, the
normalization of the Power spectrum $n=0.95$ and the amplitude of mass
density fluctuation (at redshift z$=$0) $\sigma_8=0.82$.  These
cosmological parameters are consistent with the nine-year Wilikinson
Microwave Anisotropy Probe (WMAP) results \citep{hinshaw_etal13}. A
detailed presentation of the simulation can be found in \citet{2011ApJ...740..102K}.

This results in a mass resolution of $1.35 \times 10^8$
M$_{\odot}$h$^{-1}$ for each computational particle. The completeness
limit in this simulation is set for halos with $100$ particles
corresponding to a mass of $1.35\times10^{10}$\hMsun\ or a maximum
circular velocity $V_{c}$ of $50$\kms.

We use DM halo catalogs constructed using the Bound Density Maxima (BDM)
algorithm \citep{BDM,BDMb}. To define the radius of a halo we use a
density threshold of 360 times the mean density of the Universe (
200 times the critical density), for different redshifts we use the
overdensity criterion by \cite{Bryan1998}. An important feature of BDM
is that it allows us to detect sub-halos inside larger virialized
structures.

All the raw data used in this Letter is available through the Multidark database \footnote{\texttt{www.multidark.org}}
\citep{Riebe2013}.  Furthermore, in order to facilitate the
reproducibility and reuse of our results we have made available all
the data and the source code available in a public
repository \footnote{\texttt{https://github.com/Fernandez-Trincado/Bullet\_Groups-2014}}.

To construct our main halo sample we follow three
steps. Firstly, we select all the host halos (i.e. halos that are not
inside a larger halo) with circular velocities $V_{\rm c}\geq 300$\kms
($\geq 1\times 10^{13} \hMsun$). Secondly, we select the sub-halos with
circular velocities $V_{\rm c}\geq 75$\kms ($\geq 5\times
10^{10}\hMsun$). Thirdly, we associate each host halo to its most
massive sub-halo.

Each one of the pairs (host \& sub-halo) is considered as a
potential Bullet system and is kept for the analysis described in the
next Section. We use the sub-halos as the tracer of the sub-dominant
dark matter clump in the merging cluster, i.e. the bullets.

Finally, we split the sample into two  populations; groups with
$300\kms < V_{\rm c, host}<700\kms$ and clusters with $V_{\rm
  c,host}>700\kms$. These two samples are constructed at four redshifts:
$z=0.0, 0.25, 0.5$ and $1.0$.  The number of pairs at each redshift in
each population is listed on the first arrow of Table
\ref{table:numbers}.

\ref{table:numbers}.

\section{Bullet Geometry and Measurement Setup}
\label{sec:setup}

Bullet-like configurations have two dominand dark matter structures:
the host halo and the dominant sub-halo. We describe the kinematics of this configuration by the 
position and velocity vectors of the sub-halo in a frame of reference
where the main halo is at rest; thus
$\vec{v{}}=\vec{v}_{\rm sub}-\vec{v}_{\rm host}$ and
$\vec{r{}}=\vec{r}_{\rm sub}-\vec{r}_{\rm host}$, where the subscripts $host$
and $sub$ refer to the position of the minium of potential for the
host and sub-halo in the frame of reference of the simulation, respectively.

The angle between these two vectors can be quantified by, 
\begin{equation}
  \mu\equiv
  \cos(\theta)=\frac{\vec{v{}}\cdotp{}\vec{r}}{\left\|\vec{v}{}\right\|
    \left\|\vec{r}\right\|} .
 \end{equation} 
%
This encondes information about the collision, i.e. cases of
$|\mu|\approx 1$ can be considered as head-on collisions while
$|\mu|\approx 0$ describes a grazing trajectory.

The geometrical configuration can be further described by the
following quantities. The circular velocity of each component, $V_{\rm
  c,host}$ for the host and $V_{\rm c, sub}$ for the sub-halo and the
size of the host halo $R_{\rm vir}$. Another useful quantity computed
in the simulation is the distance between the minimum of potential for
the host halo and its center of mass (computed from all the particles
inside the $R_{\rm vir}$), $X_{\rm off}=||\vec{r}_{min}-\vec{r}_{cm}||/R_{\rm vir}$, which serves as a
measurement of how much the host halo is perturbed.

In this letter we work with two quantities that could be inferred
from observations of Bullet-like systems. The projected distance
between two dominant DM clumps, $d_{\rm 2D}$ and the projected distance
between the DM-baryonic clumps, $d_{\rm 2D}^{\rm bar}$.

From the simulation point of view, the first quantity can be
translated into the 2D projected values of $||\vec{r}||$ and its value
relative to the virial radius $D_{\rm   off}= ||\vec{r}||_{\rm
  2D}/R_{\rm vir}$. The second quantity, the projected DM-baryon
distance, is not directly available from a DM-only simulation but can
be estimated from the data.

We also use the physical quantities described above to discriminate
between grazing trajectories, $|\mu|<0.9$, and Bullet-like collissions
with $|\mu|\approx 1$. In the latter case we can also distinguish three
main stages in a such encounter. First, when the sub-halo crosses the
virial radius of the host halo starting a head on collision, $D_{\rm
  off} \approx 1$ and $\mu\approx -1$. Second, as the sub-halo crosses
for the first time the center of the host halo $D_{\rm off}<1.0$ and
$\mu\approx 1$. Third, as the sub-halo reaches apogee and comes back
to the center of the halo $D_{\rm off} < 1.0$ and $\mu< -1$.

\section{Results}
\label{sec:results}

\subsection{DM-DM Displacements}
\label{fig:displacement}

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{figure_1.eps}
\end{center}
\caption{
  Integrated probability distribution for the displacement between the
  center of the host halo and its dominant sub-halo at all redshifts
  $z=0$, $0.25$, $0.5$ and $1.0$. The left panel
  shows the results in terms of the physical displacements while the
  right panel shows the displacements normalized by the virial radius
  of the host halo. The continuous (dashed) line corresponds to the halos in the
  cluster (group) sample.
  The vertical lines show the mean value and uncertainties ($186\pm
  30\hkpc$) in the  separation between the two dark matter clumps
  estimated in \citet{Gastaldello} for the \bullg. Between $40\%$
  to $60\%$ of the groups show a displacement equal or larger than
  this observational benchmark. This fraction rises to $60\%$ and
  $80\%$ in clusters.}
\label{fig:displacement}
\end{figure*}

Figure \ref{fig:displacement} presents the integrated
probability distribution for the DM-DM displacements, $d_{\rm 2D}$ The
left panel shows the displacement in physical units and the right
panel as a fraction of the virial radius of the host halo.  The panel
with the projected 2D physical displacements also shows a vertical
stripe with the estimated displacement for the Bullet-group reported
by \cite{Gastaldello}.

In the group sample we see that a fraction of
$40\%$ to $60\%$  should present a displacement equal than the
estimate for \bullg; in the cluster sample this fraction increases to
$70\%$-$80\%$. The uncertainties in this estimates are derived from
the uncertainties in the displacement measurement for \bullg. This
fraction is naturally higher in more massive systems because they are
larger in size. Normalizing the displacements by the virial radius,
right panel Figure \ref{fig:displacement}, we see that the
distributions are similar for the two samples at all redshifts.

\subsection{Collision Geometries}
\label{sec:geometry}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{figure_2.eps}
\end{center}
\caption{2D histograms in the plane $\mu$-$D_{\rm off}$. The left
  panel corresponds to clusters and the right panel to groups. The data
  used to construct this histogram includes the halos at all redshifts.}
\label{fig:geometry}
\end{figure*}

Figure \ref{fig:geometry} presents the geometry of the intereactions using the 
variables $\mu$ and $D_{\rm off}$. The first evident feature is that
most of the configurations have head-on encouters, $|\mu|>0.9$
($\theta\leq 30^{\circ}$), while only a minority with $|\mu|<0.9$ can
be described as having  grazing trajectories.

For pairs on radial trajectories there are three regions of interest
in this plane that correspond to the three merging stages described
at the end of Section \ref{sec:geometry} assuming that the sub-halo
merges (or falls below the BDM detection threshold) right at its
second pass through the center of the host halo \citep{Poole2006}.

The first region has $\mu\approx-1$ and $D_{\rm off}>0.6$, which
locates the systems where a head-on collision has just started. The
second region has $\mu\approx 1$ and $D_{\rm off}<0.6$; at this stage
the collision continues after the first crossing of the host's center,
the low number of halos with radial infalling velocities and
displacements $D_{\rm off}>0.6$ suggest that this the maximum range of
radii for the apogee.  The third region corresponds to $\mu\approx-1$
and $D_{\rm off}<0.6$ that is the the secondary infall after apogee.

In the next subsection we use the information in this sequence to
estimate the expected DM-baryon displacement.

\subsection{DM-baryon Displacement}
\label{sec:baryonic_displacements}

The results we have derived so far apply to multi-modal systems and
the expected separation between the two dominant dark matter
clumps. However, a non-zero DM-DM displacement does not imply 
a non-zero DM-baryon displacement. Given that we use a DM only
simulation, we have to find a way to estimate a DM-baryon displacement
from the available information.

This estimate is based on the different kinds of trajectories and
collision stages described in previous sections. To start with, we
consider that systems with $|\mu|<0.9$, describing grazing
trajectories, have a baryonic displacement equal to zero.  In systems
with head-on collissions, $|\mu|>0.9$, we distinguish two different
cases. The first one, systems with  $\mu<-0.9$ and large displacements $D_{\rm
  off}>0.6$ most probably describe the beginning of the interaction
and should have baryonic displacements equal to zero.
equal to zero. In all the other cases, the DM-baryon displacement
should be equal to zero. We approximate this displacement by  $d_{\rm
  2D}^{\rm bar} = X_{\rm off} R_{\rm   vir}$, where $X_{\rm   off}$ is
the offset computed between the minium of potential and the center of
mass for each host halo computed from all the matter inside the virial
radius.

This simplified model does not take into account that there is a
fraction of halos with $\mu<-0.9$ and $D_{\rm off}<0.6$ for which the
collision has not started and should have $d_{2D}^{\rm bar}=0$. A detailed
modeling of this fraction requires the study of the complete merger
tree of the halo and sub-halo, a study beyond the scope of this
Letter. Instead we caution the reader that the derived fraction of
halos with a displacement $<d_{2D}^{\rm bar}$ must be considered as an
upper limit.

The results for the integrated distributions for $>d_{\rm 2D}^{\rm bar}$
are shown in Figure \ref{fig:baryonic_displacements}. The dashed lines
represented the results for groups and the continuous lines correspond
to clusters. As a test of our model we compare the cluster results
against the analytic fit provided by \cite{ForeroRomero2010}. This fit
reproduces the statistics for the DM-baryon separation found for
clusters more massive than $>10^{14}$\hMsun\ in a simulation
which included a description for DM and gas with $8$ times more volume
and $8$ times less mass resolution than the Bolshoi Simulation. The
fit is valid for separations larger than $70$\hkpc, beyond which we
find that it provides a remarkably good description within a factor of
$\sim 2$ of our results. This gives us confidence in our approach to
estimate the expected fraction of groups with a DM-baryon
displacement.

Finally, from Figure \ref{fig:baryonic_displacements} we see that only a
fraction of $0.1\%$ to $1\%$ of the groups are expected to have a
DM-baryon displacement equal or larger than $87\pm 14$\hkpc as
observed in \bullg. This fraction rises to $4\%$ to $10\%$ in the case
of clusters, consistent with the results reported by
\cite{ForeroRomero2010}.

\subsection{Towards a Statistical Comparison Against Observations}

Recently \cite{Foex2013} presented an analysis of $80$ galaxy groups
in the SL2S sample. From the light distribution, only $34$ objects
($\sim 42\%$) have regular isophotes, $33$ had ellongated
isophotes (hints of merging system), and $13$ ($\sim 16\%$) had a
clear bi-modal light distribution; \bullg\ is one of these
$13$ systems.

The bi-modal objects are defined to have at least a clear second
luminosity peak within $350$\hkpc from the main halo, as traced by the
strong lensing system.  The lowest separation in those systems is
$64$ \hkpc and the average is $145\pm 52$\hkpc.

We now make a comparison of these fractions against the results of our
simulations. The results are summarized in Table
\ref{table:numbers}. The first row indicates the total number of halos
in each sample at each redshift. The second row shows the number of
objects with DM-DM displacements $46\hkpc< d_{\rm 2D}< 350$\hkpc. The
last row indicates the number of objects in the previous sub-sample
with DM-baryon displacements $d_{\rm 2D}^{\rm bar}>87$\hkpc.

Considering that the statistics for the cluster sample are dominated
by objects in the mass range of the \bullg we make a comparison
against this sample. Table \ref{table:numbers} shows that $\sim 40\%$ of the 
clusters are expected to have large DM-DM displacements, this is a
factor of $\sim 2$ larger than the observational estimate by
\cite{Foex2013}. However, roughly $\sim 7\%$ of these systems present a
displacement equal or larger than the observed in the \bullg, a
fraction that is compatible with the $1/13\sim 0.07$ fraction in the
SL2S sample from which \bullg\ was drawn.

This rough comparison  shows that our estimates for the relative number of
bullet systems (large DM-baryon displacements) with respect to
multi-modal systems (large DM-DM displacements) is compatible with
observations. A proper comparison must take into account all the observational
uncertainties, biases, mixes between our two populations (groups \&
clusters at different redshifts) to derive a stronger bound from
the simulation, something that is beyond the scope of this Letter.

Neverthless, comparing the number for group and cluster samples in the
last row of Table \ref{table:numbers} we can already predict that a
fixed redshift the number density of bullets similar to \bullg that
three times larger for groups than for clusters.

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{figure_3.eps}
\end{center}
\caption{Integrated probability distribution for the estimated
  baryonic displacements in the group and cluster samples. Continous (dashed)
  lines correspond to clusters (groups). The continuous black line
  marked as $P_{2D}$ shows the statistics reported by
  \citet{ForeroRomero2010} for a cosmological simulation including DM
  and baryons. The vertical lines correspond to the mean value and
  uncertainty of the displacement measured for \bullg.} 
\label{fig:baryonic_displacements}
\end{figure*}

\begin{table*}
\begin{center}
\begin{tabular}{ccccccccc}\hline\hline
        & Groups & Groups & Groups & Groups & Clusters & Clusters & Clusters & Clusters\\
  & $z=0$   & $z=0.25$   & $z=0.5$   & $z=1.0$   & $z=0$   & $z=0.25$   & $z=0.5$   & $z=1.0$ \\
  & ($\# $)    & ($\# $)    & ($\# $)    & ($\# $)    & ($\# $)    & ($\# $)    & ($\# $)    & ($\# $)  \\\hline
Full Sample & 9641 & 9984 & 10244 & 10190 & 400  & 363 & 310 & 192 \\
$64 <d_{\rm 2D}/\hkpc< 350$ & 6188 & 6422 & 6635 & 6933 & 151 &
141 & 120 & 99\\
($64 <d_{\rm 2D}/\hkpc< 350 $) \& ($d_{\rm 2D}^{\rm bar}/\hkpc >87$) & 14 & 25 & 44 & 35 & 8 & 9 & 13 & 8 \\\hline\hline
\end{tabular}
\caption{Absolute number of objects in the groups and clusters sample
  at all redshifts for different selection criteria. }
\label{table:numbers}
\end{center}
\end{table*}

\section{Conclusions}
\label{sec:conclusions}

In this Letter we estimate the fraction of galaxy groups and clusters
in a $\Lambda$CDM cosmology that could present observational features
associated to a bullet-like event. This is motivated by the recent
observational results of \citep{Gastaldello} where a system (\bullg)
on the mass range $1\times 10^{14}$\hMsun\ and velocity dispersion
$650$\kms was reported to feature a displacement between its baryonic
(hot gas) and dark matter components.

We compute the distribution of projected displacements
between the dominant DM clumps in two kinds of systems; groups with
circular velocities $300\kms<V_{\rm c}<700\kms$ ($1.0\times
10^{13}\hMsun < M_{\rm vir}<1.0\times 10^{14}\hMsun$) and clusters with $V_{\rm
  c}>700$\kms ($M_{\rm vir}>1.0\times 10^{14}\hMsun$). We report these
results at four different redshifts $z=0.0,0.25,0.5$ and $10$. Our
results are based on large DM-only N-body cosmological simulation with
a resolution that allows to study for the first time Bullet-like
configurations in the mass range of galaxy groups.

Our main result is that a fraction of $40\%$-$60\%$ of the halos
in the group sample present DM-DM displacement equal or larger than the
observed displacement for \bullg. For halos in the cluster sample this
fraction increases to $60\%$-$80\%$. We also derive an
estimate for the DM-baryon displacement In the group sample
$0.1\%$-$1.0\%$ of the halos show a displacement equal or larger than
the  measurements of \bullg\ by \citep{Gastaldello}; in the cluster
sample this fraction rises to $4\%$-$10\%$.

For the case of \bullg\ a fair comparison is achieved against our cluster
sample which has statistics dominated by objects of similar mass. In
a rough comparison using the observational criteria \citep{Foex2013,
  Gastaldello} we find that the relative number of bullets
(large DM-baryon displacement) with respect to a general sample of
multi-modal systems (large DM-DM displacements) is consistent with
observations; both are in the range $\sim 7\%$.

In terms of absolute numer densities, we expect to have $\sim 6.0\times
10^{-7}$ Mpc$^{-3}$ groups similar to \bullg; three times the expected
value for clusters. This opens up a new observational possibility with surveys
such as SL2S that target a large number of groups and estimate its
multi-modal nature from lensing analysis \citep{Foex2013}. An approach
that can be further exploited with upcoming lensing surveys (e.g. with
the {\it Euclid} satellite).

The CosmoSim database used in this paper is a service by the
Leibniz-Institute for Astrophysics Potsdam (AIP). The  Bolshoi
simulation was performed within the Bolshoi project of the University
of California High-Performance AstroComputing Center (UC-HIPACC) and
was run at the NASA Ames Research Center.

J.E.F-R acknowledges support from Vicerrector\'ia de
Investigaciones through a FAPA starting grant.

T.V. acknowledges support from CONACYT through grant 165365 and
203489 through the program Estancias posdoctorales y sab\'aticas al
extranjero para la consolidaci\'on de grupos de investigaci\'on.

V.M. gratefully acknowledges FONDECYT support through grant 1120741
and ECOS/CONICYT through grant C12U02.

\bibliographystyle{apj}
\bibliography{references}

\end{document}